HMM Model development

### 2020-05-20 dev/hmm_model_devel.Rmd
More encapsulation, fitting to Grein et al. dataset

```{r}
data <- hmm_simulator(20, 15, 2, use_noisy_states = FALSE, N_treatments = 0)

@@

```{r}
data_grein <- readxl::read_excel(here::here("public_data", "Grein_et_al.xlsx"), na = c("")) %>%
  select(-Remdesivir_Last_Day) %>%
  pivot_longer(starts_with("Day_"), names_prefix = "Day_", names_to = "day", values_to = "breathing") %>%
  filter(!is.na(breathing)) %>%
  transmute(
    patient = Patient,
    breathing = breathing,
    day = as.integer(day)) %>%
  group_by(patient) %>%
  mutate(breathing = if_else(breathing == "ND", breathing[which.max(day[breathing %in% c("AA", "LoO2", "HiO2", "NIPPV", "MV", "ECMO")])], breathing)) %>%
  ungroup() %>%
  mutate(
    breathing = factor(breathing, levels = c("Discharge", "AA", "LoO2", "HiO2", "NIPPV", "MV", "ECMO", "Death"), ordered = TRUE)
  )

@@

@@

>>>>

```{r}
data <- hmm_simulator(20, 15, 2, use_noisy_states = TRUE, N_treatments = 0)

@@

disease_levels <- c("AA", "LoO2", "HiO2", "NIPPV", "MV", "ECMO")
breathing_levels <- c("Discharge", disease_levels, "Death")

data_grein <- readxl::read_excel(here::here("public_data", "Grein_et_al.xlsx"), na = c("")) %>%
  select(-Remdesivir_Last_Day) %>%
  pivot_longer(starts_with("Day_"), names_prefix = "Day_", names_to = "day", values_to = "breathing") %>%
  filter(!is.na(breathing)) %>%
  transmute(
    patient = Patient,
    breathing = breathing,
    day = as.integer(day)) %>%
  group_by(patient) %>%
  mutate(breathing_imputed = if_else(breathing == "ND", breathing[which.max(day[breathing %in% c("AA", "LoO2", "HiO2", "NIPPV", "MV", "ECMO")])], breathing)) %>%
  ungroup() %>%
  mutate(
    breathing = factor(breathing, levels = breathing_levels, ordered = TRUE),
    breathing_imputed = factor(breathing, levels = breathing_levels, ordered = TRUE)
  )

@@

series_data <- data_grein %>% transmute(.serie = as.integer(patient), .time = day + 1, .observed = breathing)

hidden_state_data <- data.frame(id = factor(breathing_levels, levels = breathing_levels)) %>% mutate( corresponding_obs = id)

rate_data <- rbind(
  data.frame(.from = disease_levels, .to = "Death", rate_group = "death"),
  data.frame(.from = "AA", .to = "Discharge", rate_group = "improve_one"),
  data.frame(.from = disease_levels[2 : length(disease_levels)], .to = disease_levels[1 : (length(disease_levels) - 1)], rate_group = "improve_one"),
  data.frame(.from = disease_levels[1 : (length(disease_levels) - 1)], .to = disease_levels[2 : length(disease_levels)], rate_group = "worsen_one")
) %>%
  mutate(.from = factor(.from, levels = breathing_levels),
         .to = factor(.to, levels = breathing_levels),
         )

initial_states <- series_data %>% filter(.time == 1) %>% arrange(.serie) %>% pull(.observed)

prior = brms::set_prior("normal(-2, 5)", "Intercept") +
  brms::set_prior("normal(0, 3)", "b") +
  brms::set_prior("normal(0, 1)", "sd", group = ".rate_id")

grein_hmm <- brmshmmdata(~ rate_group + (1|.rate_id), series_data, rate_data, hidden_state_data, initial_states, prior = prior)

fit_grein <- brmhmm(grein_hmm)

@@

```{r}
# fit_grein$data <- grein_hmm
# fit_grein$data_processed <- make_data_hmm(grein_hmm)

#posterior_grein <- brms::posterior_epred(fit_grein)
posterior_grein <- brms::posterior_predict(fit_grein)

posterior_grein_wide <- prediction_to_wide_format(fit_grein$data, posterior_grein)

posterior_grein_wide <- posterior_grein_wide %>% mutate(.predicted = factor(.predicted, levels = 1:length(levels(fit_grein$data$serie_data$.observed)), ordered = TRUE, labels = levels(fit_grein$data$serie_data$.observed)))

step_size <- 6
for(step in 1:ceiling(fit_grein$data_processed$standata$N_series / step_size)) {
  series_id <- ((step - 1) * step_size + 1) : (step* step_size)

  posterior_state_plot(posterior_grein_wide, series_id, fit_grein$data) %>% print()

}

#TODO ppchecks: last state, pairwise state transitions, step up/down

```{r}
d1 <- data.frame(x = 1, y = factor(1:3, levels = 1:3, labels = c("First","After first","After after first"), ordered = TRUE), f = c(0.1,0.2,0.3))

d2 <- data.frame(x = c(1,2), y2 = factor(c(2,3), levels = 1:3, labels = c("First","After first","After after first"), ordered = TRUE), f = 0.5)

ggplot(d1, aes(x = x, y = y, fill = f)) + geom_raster() + 
  geom_line(data = d2, inherit.aes = FALSE, aes(x = x, y = y2, group = 1))
